Global code motion of congruent instructions

### 2017-04-17 paper.tex
anonymous and change title
fix section 3

%% For double-blind review submission
\documentclass[acmlarge,review]{acmart}\settopmatter{printfolios=true}
%\documentclass[acmlarge,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission
%\documentclass[acmlarge,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission
%\documentclass[acmlarge]{acmart}\settopmatter{}

@@

%% Title information
\title[SSA GCM]{SSA based global code motion of identical computations}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

@@

The original algorithm presented in \cite{das2012} has mismatched types in terms
of uses and nodes because each node can have many instructions and hence many
uses. Also, while iterating on the dominator of each user of a variable we may
reach to the beginning of a function, in that case the inner while loop needs to
terminate. These two cases were missing from the algorithm and we came across
them during implementation. If a variable is not live-out of a basic block, it
still may be used later in the basic block. To establish whether a use of a
variable is a kill we iterate on all the subsequent instructions in a basic
block checking for uses should isLiveOutUsingMergeSet return false.

@@

The current implementation of \GVN{} in \LLVM{} has some limitations when it
comes to loads and stores so we compute the \GVN{} of loads and stores
separately.  Our solution to value number loads is to hash the address from
where the value is to be loaded. For stores, we value number the address as well
as the value to be stored at that address. Another limitation of the current
\GVN{} implementation in \LLVM{} is that the instructions dependent on the loads
will not get numbered correctly, and so after hoisting all candidates we need to
rerun the \GVN{} analysis in order to discover new candidates now available
after having hoisted load instructions.  This limitation should be addressed in
a new implementation of the \GVN{} based on \MemorySSA{}, that would better
account for equivalent loads and their dependent instructions.

@@

\subsection{Legality check}
\label{subsec:legality}
Since the equality of candidates is purely based on the value numbers, we also
need to establish if hoisting them to a common dominator or sinking them to a
common post-dominator would be legal. Once a common dominator (post-dominator)
is found, we check whether all the use-operands of the set of instructions are
available at that position. It is possible reinstantiate (remateralize) the
use-operands in some cases when the operands are not available and make it legal
to move the instruction.

Subsequently, it is checked that the side-effects of the computations (if any)
do not intersect with any side-effects between the instructions to be
hoisted/sunk and their hoisting/sinking point. It is also necessary to check if
there are indirect branch targets e.g., landing pad, case statements, goto
labels etc., along the path because it becomes difficult to prove safety checks
in those cases. In our current implementation we discard candidates on those
paths.

@@

\subsubsection{Legality of sinking expressions}
Sometimes, hoisting is not upward-safe \ref{click1995global} e.g., if the
expressions are in a landing pad etc., in that case sinking of those expressions
may reduce code size. Sinking may also reduce the register pressure in some
cases e.g., when the use operands are not kills.For sinking, higher ranked
expressions would be sunk first. And it would be illegal to sink higher ranked
identical expressions if they are not anticipable in the common
post-dominator. For example:

>>>>

%% For double-blind review submission
\documentclass[acmlarge,review,anonymous]{acmart}\settopmatter{printfolios=true}
%\documentclass[acmlarge,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission
%\documentclass[acmlarge,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission
%\documentclass[acmlarge]{acmart}\settopmatter{}

@@

%% Title information
\title[SSA GCM]{Global code motion of congruent instructions}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
%\subtitle{Subtitle}                     %% \subtitle is optional
%\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

@@

The original algorithm presented in \cite{das2012} has mismatched types in terms
of uses and nodes because each node can have many instructions and hence many
uses. Also, while iterating on the dominator of each user of a variable we may
reach to the beginning of a function, in that case the inner while loop needs to
terminate. These two cases were missing from the algorithm and we came across
them during implementation. If a variable is not live-out of a basic block, it
still may be used later in the basic block: to establish whether a use of a
variable is a kill we iterate on all the subsequent instructions in a basic
block checking for uses.

@@

The current implementation of \GVN{} in \LLVM{} has some limitations when it
comes to loads and stores so we compute the \GVN{} of loads and stores
separately.  Our solution is to value number the address from where the value of
a load is to be read from memory. For stores, we value number the address as
well as the value to be stored at that address. Another limitation of the
current \GVN{} implementation in \LLVM{} is that the instructions dependent on
the loads will not get numbered correctly, and so after hoisting all candidates
we need to rerun the \GVN{} analysis in order to discover new candidates now
available after having hoisted load instructions.  This limitation should be
addressed in a new implementation of the \GVN{} based on \MemorySSA{}, that
would better account for equivalent loads and their dependent instructions.

@@

\subsection{Legality check}
\label{subsec:legality}
Since the equality of candidates is purely based on the value numbers, we also
need to establish if hoisting them to a common dominator or sinking them to a
common post-dominator would be legal. Once a common dominator (post-dominator)
is found, we check whether all the use-operands of the set of instructions are
available at that position. It is possible to reinstantiate or remateralize the
use-operands in some cases when the operands are not available and make it legal
to move the instruction.

Subsequently, it is checked that the side-effects of the computations (if any)
do not intersect with any side-effects between the instructions to be
hoisted/sunk and their hoisting/sinking point. It is also necessary to check if
there are indirect branch targets, e.g., landing pad, case statements, goto
labels etc., along the path because it becomes difficult to prove safety checks
in those cases. In our current implementation we discard candidates on those
paths.

@@

\subsubsection{Legality of sinking expressions}
Sometimes, hoisting is not upward-safe \cite{click1995global}, e.g., if the
expressions are in a landing pad, sinking of those expressions
may reduce code size. Sinking may also reduce the register pressure in some
cases, e.g., when the use operands are not kills. For sinking, higher ranked
expressions would be sunk first. And it would be illegal to sink higher ranked
identical expressions if they are not anticipable in the common
post-dominator. For example:
